### 2021-09-28 revision/ch6conc.tex
done

This thesis has described bedload transport from the statistical physics of individual grains.
The work has improved upon earlier descriptions.
It describes the movements of grains along a wider range of timescales than before, highlights Langevin equations, master equations, and random walks as the common threads of stochastic sediment transport modeling, and provides more realistic descriptions of bedload transport. In particular, I have
\begin{enumerate}
	\item developed the linkage between sediment flux probability distributions and individual particle trajectories (Ch. \ref{ch:flux}), 
	\item described particle trajectories alternating through motion and rest with fluctuating velocities (Ch. \ref{ch:flux}),
	\item evaluated the control of bed elevation changes over sediment transport rates (Ch. \ref{ch:ch3}),
	\item characterized how long particles can remain buried within the sedimentary bed (Ch. \ref{ch:ch3}),
	\item incorporated the process of sediment burial in a description of downstream particle movement (Ch. \ref{ch:downDiff}), and
	\item formulated the velocity distributions of sediment particles including episodic particle-bed collisions (Ch. \ref{ch:langevin}).
\end{enumerate}
These developments bring us closer to a comprehensive mathematical description of sediment transport in streams.

@@

Once the stochastic dynamical equations were written for a given problem, its solutions were averaged over realizations of the control variables to produce a master equation, an integro-differential equation governing the probability distributions of the response variables.

@@

\subsection{Calculation of the probability distribution of the sediment flux from micro-mechanics of particle transport}

@@

This description predicts multiple-range diffusion across local, intermediate, and global scales as a result of particle velocity fluctuations within the motion state; it introduces a dimensionless P\'{e}clet number as an important characteristic of bedload sediment transport; and it relates the timescales at which transitions between local, intermediate, and global timescales occur to the movement characteristics of individual grains.
These developments constitute a new understanding of bedload movement across its timescales and invite generalizations aimed at pushing the approach further to include interactions between particles and temporal correlations in bedload movement velocities.

@@

This work demonstrates that this negative feedback buffers bedload transport fluctuations whenever collective entrainment occurs, meaning the magnitude of bedload transport fluctuations depends on the relative bed elevation. The residence times of buried particles are random variables that lie on heavy-tailed power-law distributions. These distributions allow for arbitrarily long resting times, which poses challenging implications for researchers attempting to determine the downstream sediment flux for applications by tracking sediment tracer particles.

\subsection{Characterization of how sediment burial affects the downstream transport of sediment particles}

@@

Until this work, the mechanisms which produce the different spreading rates of particles across the scaling ranges identified by Nikora and coworkers had been uncertain for several decades, and earlier works had included what were believed to be the required features without describing three or more scaling ranges.
Ch. \ref{ch:downDiff} brings some closure to these studies and demonstrates that many approaches to describe individual particle motions are reformulations of the continuous time random walk formalism from physics, indicating unity within a diverse body of research and bringing powerful tools from statistical physics to the sediment transport problem.

@@

The stochastic descriptions of sediment transport developed in this thesis show a methodology we can use to step beyond averaged descriptions of landscape evolution, propagate noises through the equations governing landscape change, and revisit the old question in geomorphology: How does variability shape Earth's surface?

>>>>

This thesis has described bedload transport from the statistical physics of individual grains.
This work describes the movements of grains along a wider range of timescales than before, highlights Langevin equations, master equations, and random walks as the common threads of stochastic sediment transport modeling, and provides more realistic descriptions of bedload transport. In particular, within the thesis I have
\begin{enumerate}
	\item developed the linkage between sediment flux probability distributions and individual particle trajectories (Ch. \ref{ch:flux}), 
	\item described particle trajectories alternating through motion and rest with fluctuating velocities (Ch. \ref{ch:flux}),
	\item evaluated the control of bed elevation changes over sediment transport rates (Ch. \ref{ch:ch3}),
	\item characterized how long particles can remain buried within the sedimentary bed (Ch. \ref{ch:ch3}),
	\item incorporated the process of sediment burial in a description of downstream particle movement (Ch. \ref{ch:downDiff}), and
	\item formulated the velocity distributions of sediment particles including episodic particle-bed collisions (Ch. \ref{ch:langevin}).
\end{enumerate}
These developments bring us closer to a comprehensive mathematical description of sediment transport in streams and bring powerful physics and mathematics tools to longstanding problems in river science.

@@

Once the stochastic dynamical equations were written for a given problem, its solutions were averaged over realizations of the control variables to produce a master equation for the problem, an integro-differential equation governing the probability distributions of the response variables.

@@

\subsection{Calculation of the probability distribution of the sediment flux from individual particle trajectories}

@@

This description predicts multiple-range diffusion across local, intermediate, and global scales as a result of particle velocity fluctuations within the motion state; it introduces a dimensionless P\'{e}clet number as an important characteristic of bedload sediment transport; and it relates the timescales at which transitions between local, intermediate, and global timescales occur to the movement characteristics of individual grains.
These developments constitute new understanding of bedload movement across its timescales and invite generalizations that include interactions between particles and temporal correlations to bedload movement velocities.

@@

This work demonstrates that this negative feedback buffers bedload transport fluctuations whenever collective entrainment occurs, meaning the magnitude of bedload transport fluctuations depends on the relative bed elevation. The residence times of buried particles are random variables that lie on heavy-tailed power-law distributions. These distributions allow for arbitrarily long resting times, which poses challenging implications for researchers attempting to determine downstream sediment fluxes from sediment tracer data.

\subsection{Characterization of sediment burial controls on the downstream transport of sediment particles}

@@

Until this work, the mechanisms which produce the different spreading rates of particles across the scaling ranges identified by Nikora and coworkers had been uncertain for several decades, and earlier works had included what were believed to be the required features without describing three or more scaling ranges.
Ch. \ref{ch:downDiff} brings some closure to these studies and demonstrates that many approaches to describe individual particle motions are particular adaptations of the continuous time random walk formalism from physics, indicating unity within a diverse body of research and bringing powerful tools from statistical physics to the sediment transport problem.

@@

The stochastic descriptions of sediment transport developed in this thesis show a methodology we can use to step beyond averaged descriptions of geomorphic change, propagate noises through the equations governing landscape change, and revisit the old question in geomorphology: How does variability shape Earth's surface?

### 2021-09-28 revision/laySum.tex
done

Predicting the flow rates of gravel through river channels is important to manage environmental issues involving rivers.
Most approaches to predict gravel flow consider gravel as if it were a heavy fluid.
This description oversimplifies the problem, since gravel flow is actually individual grains bouncing and rolling downstream, which is hardly like a fluid.

@@

\endinput
The lay or public summary explains the key goals and contributions of
the research\slash{}scholarly work in terms that can be understood by the
general public. It must not exceed 150 words in length.

>>>>

Predicting the flow rates of gravel through river channels is important to manage environmental issues involving rivers.
Most approaches to predict gravel flow consider gravel as if it were a heavy fluid.
These descriptions oversimplify the problem, since gravel flow is actually composed of individual grains bouncing and rolling downstream, which is hardly like a fluid.

@@

### 2021-09-28 revision/pref.tex
done

This thesis is entirely the original research of Kevin Pierce, including all figures, writing, and calculations.
The supervisory committee and especially the research supervisor Marwan Hassan provided guidance in the research.
Suggestions from Michael Church, Rui Ferreira, Roger Beckie, and Peter Ashmore improved the clarity of the writing.

The thesis includes two published works, Chapters \ref{ch:ch3} and \ref{ch:downDiff}, available in \textit{Journal of Geophysical Research: Earth Surface} and \textit{Geophysical Research Letters} respectively, with Marwan Hassan as coauthor \citep{Pierce2020a,Pierce2020b}. Chapters \ref{ch:flux} and \ref{ch:langevin} will be submitted for publication after acceptance of the thesis.

>>>>

This thesis is entirely the original research of Kevin Pierce, including all figures, writing, and calculations.
The supervisory committee and especially the research supervisor Marwan Hassan provided guidance in the research.
Suggestions from Michael Church, Rui Ferreira, Roger Beckie, Peter Ashmore, Bernard Laval, and Brett Eaton improved the clarity of the writing.

The thesis includes two published works, Chapters \ref{ch:ch3} and \ref{ch:downDiff}, available in \textit{Journal of Geophysical Research: Earth Surface} and \textit{Geophysical Research Letters} respectively, with Marwan Hassan as coauthor \citep{Pierce2020a,Pierce2020b}. Chapters \ref{ch:flux} and \ref{ch:langevin} will be submitted for publication following acceptance of the thesis.
